Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-27 filter.tex
Updates to Sections 3 & 4.
Edits to filtering section.

The instruction-grained monitoring techniques we consider forward all
instructions of certain types (e.g., load instructions) to the monitoring core.
However, not all of these forwarded instructions have relevant monitoring to be
done. For example, for array bounds check, all load and store instructions are
forwarded. However, only loads and stores associated with array pointers have
relevant monitoring operations to be done. Thus, we can greatly reduce the amount of
monitoring done by only forwarding instructions to the monitoring core that 
correspond to array pointers. These accesses can be identified because their
corresponding metadata has been set with base and bound information.
More generally, we can reduce monitoring overheads by filtering out 
events that correspond to clean or uninitialized metadata.

@@

\subsection{Tracking Clean Metadata}

@@

Figure~\ref{fig:filter.architecture} shows a block diagram of the dataflow
tracking engine for
keeping track of clean metadata. Initially, all metadata is marked as clean.
When the monitoring core writes a non-null metadata value, it also sends a
signal to the dataflow engine to mark the metadata as dirty.

When a monitoring event arrives at the dataflow engine,
it reads in up to two clean flags. A table is used to configure
which clean flags should be read in depending on the instruction type of
the event and information from the monitoring event. A pair of simple ALUs are
used to allow calculating memory addresses based on information from the
monitoring event. Typically, the clean flags read correspond to the
source operands of the monitored instruction.

Given this set of clean flags, a Filter Decision Table decides, based
on instruction type, whether an event can be filtered out. This table is
configured based on the specific semantics of a monitoring technique.
Typically, if all clean flags are marked as clean, then the
event can be filtered out. For example, for array bounds check, if the source
register's metadata is marked as clean, then this means that the
register does not correspond to an array pointer. Thus, we can filter out the
monitoring operation for this event. 
It is important that we also propagate the clean information to the destination
of the instruction.  The Filter Decision Table also includes a bit indicating
whether the destination operand metadata should be marked as uninitialized.

>>>>

For the instruction-grained monitoring techniques we consider, instructions in
the main program's execution are forwarded to the monitoring core based on
their instruction type (e.g., load, store, ALU, etc.).
However, not all of these forwarded instructions have relevant monitoring to be
done. For example, for array bounds check, all load and store instructions are
forwarded. However, only loads and stores associated with array pointers have
relevant monitoring operations to be done. Thus, we can greatly reduce the amount of
monitoring done by only forwarding instructions to the monitoring core that 
correspond to array pointers. These accesses can be identified because their
corresponding metadata has been initialized with base and bound address information.
More generally, we can reduce monitoring overheads by filtering out 
events that correspond to clean or uninitialized metadata.

@@

\subsection{Dataflow Engine for Clean Metadata}

@@

Figure~\ref{fig:filter.architecture} shows a block diagram of the dataflow
tracking engine for
keeping track of clean metadata. The engine keeps a \emph{clean flag}
corresponding to each piece of register and memory metadata. Initially, all
metadata is marked as clean.
When the monitoring core writes a non-null metadata value, it also sends a
signal to the dataflow engine to mark the metadata as dirty (i.e., initialized).

When a monitoring event arrives at the dataflow engine,
it reads in up to two clean flags. These typically correspond to the source
operands of the monitored instruction but can be configured to other targets. 
This is done by configuring a pair of address generation units during the
initialization of the monitoring core.
For example, this can be used to keep clean flags at the word-level rather than
the byte-level by shifting memory addresses.

Given this pair of clean flags, a filter decision table decides, based
on instruction type, whether an event can be filtered out. This table is
configured based on the specific semantics of a monitoring technique.
Typically, if all clean flags are marked as clean, then the
event can be filtered out. For example, for array bounds check, if the source
register's metadata is marked as clean, then this means that the
register does not correspond to an array pointer. Thus, we can filter out the
monitoring operation for this event.

It is important that we also propagate the clean information to the destination
of the instruction in order to prevent the use of stale (possibly non-clean) metadata.
The filter decision table also indicates
whether the destination operand metadata should be marked as clean.
